---
title: "Rmarkdown_LM"
author: "Leah Moubadder, Jasmine Aqua, Susan Hoffman"
date: "11/15/2021"
output: pdf_document
---
---
title: "Spatial Epi Project"
author: "Leah Moubadder, Jasmine Aqua, Susan Hoffman"
date: "11/11/2021"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objective: 

# Loading packages
```{r load packages}
pacman::p_load(tidyverse, # General data wrangling
raster,
rgdal,
tidycensus,
sp,
readxl,
tigris,
spatialreg,
spdep, # Functions for creating spatial weight, spatial analysis
sf, # manage spatial data using simple features
tmap) # mapping of sf or sp data
```

#Loading data

```{r ENVIRO desc maps}
setwd("C:/Users/lmoubad/OneDrive - Emory University/spatial-epi/spatialfinal2021")


  full_data <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data")
  full_data_georgia <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_georgia")
  full_data_atlanta <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_atlanta")
  full_data_augusta <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_augusta")
  full_data_macon <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_macon")
  full_data_savannah <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_savannah")
  full_data_columbus <- readOGR(dsn=path.expand("Data/tempdir"),layer="full_data_columbus")



```
# Environmental 15 panel maps
```{r env desc maps}
MyPalette_HOLC <- c("#397A1C", "#3581D9", "#FFFF33", "#E31A1C")

MyLabels <- c("Low (Q1)", "Medium (Q2)", "High (Q3)", "Very High (Q4)")

map1 <-  tm_shape(full_data_atlanta) + 
  tm_fill('RESP',
          n = 4,
          style = 'quantile',
          palette = "Blues", 
          labels = MyLabels, 
          title = "RESP") +
  tm_layout(title = "RESP in Atlanta") +
  tm_borders()

map2 <- tm_shape(full_data_macon) + 
  tm_fill('RESP',
          n = 4,
          style = 'quantile',
          palette = "Blues", 
          labels = MyLabels, 
          title = "RESP") +
  tm_layout(title = "RESP in Macon") +
  tm_borders()
  
  map3 <- tm_shape(full_data_columbus) + 
  tm_fill('RESP',
          n = 4,
          style = 'quantile',
          palette = "Blues", 
          labels = MyLabels, 
          title = "RESP") +
  tm_layout(title = "RESP in Columbus") +
  tm_borders()
  
    map4 <- tm_shape(full_data_augusta) + 
  tm_fill('RESP',
          n = 4,
          style = 'quantile',
          palette = "Blues", 
          labels = MyLabels, 
          title = "RESP") +
  tm_layout(title = "RESP in Augusta") +
  tm_borders()
  
   
  map5 <- tm_shape(full_data_savannah) + 
  tm_fill('RESP',
          n = 4,
          style = 'quantile',
          palette = "Blues", 
          labels = MyLabels, 
          title = "RESP") +
  tm_layout(title = "RESP in Augusta") +
  tm_borders()
  
 
      
  map6 <- tm_shape(full_data_atlanta) + 
  tm_fill('DSLPM',
          n = 4,
          style = 'quantile',
          palette = "Greens", 
          labels = MyLabels, 
          title = "DSLPM") +
  tm_layout(title = "DSLPM in Atlanta") +
  tm_borders()

    map7 <- tm_shape(full_data_macon) + 
  tm_fill('DSLPM',
          n = 4,
          style = 'quantile',
          palette = "Greens", 
          labels = MyLabels, 
          title = "DSLPM") +
  tm_layout(title = "DSLPM in Macon") +
  tm_borders()
    
        map8 <- tm_shape(full_data_columbus) + 
  tm_fill('DSLPM',
          n = 4,
          style = 'quantile',
          palette = "Greens", 
          labels = MyLabels, 
          title = "DSLPM") +
  tm_layout(title = "DSLPM in Columbus") +
  tm_borders()

        map9 <- tm_shape(full_data_augusta) + 
  tm_fill('DSLPM',
          n = 4,
          style = 'quantile',
          palette = "Greens", 
          labels = MyLabels, 
          title = "DSLPM") +
  tm_layout(title = "DSLPM in Augusta") +
  tm_borders()
        
             map10 <- tm_shape(full_data_savannah) + 
  tm_fill('DSLPM',
          n = 4,
          style = 'quantile',
          palette = "Greens", 
          labels = MyLabels, 
          title = "DSLPM") +
  tm_layout(title = "DSLPM in Savannah") +
  tm_borders()
             
             
        
  map11 <- tm_shape(full_data_atlanta) + 
  tm_fill('CANCER',
          n = 4,
          style = 'quantile',
          palette = "Purples", 
          labels = MyLabels, 
          title = "CANCER") +
  tm_layout(title = "CANCER in Atlanta") +
  tm_borders()

    map12 <- tm_shape(full_data_macon) + 
  tm_fill('CANCER',
          n = 4,
          style = 'quantile',
          palette = "Purples", 
          labels = MyLabels, 
          title = "CANCER") +
  tm_layout(title = "CANCER in Macon") +
  tm_borders()
    
        map13 <- tm_shape(full_data_columbus) + 
  tm_fill('CANCER',
          n = 4,
          style = 'quantile',
          palette = "Purples", 
          labels = MyLabels, 
          title = "CANCER") +
  tm_layout(title = "CANCER in Columbus") +
  tm_borders()

        map14 <- tm_shape(full_data_augusta) + 
  tm_fill('CANCER',
          n = 4,
          style = 'quantile',
          palette = "Purples", 
          labels = MyLabels, 
          title = "CANCER") +
  tm_layout(title = "CANCER in Augusta") +
  tm_borders()
        
             map15 <- tm_shape(full_data_savannah) + 
    tm_fill('CANCER',
          n = 4,
          style = 'quantile',
          palette = "Purples", 
          labels = MyLabels, 
          title = "CANCER") +
  tm_layout(title = "CANCER in Savannah") +
  tm_borders()

              tmap_arrange(map1, map2, map3, map4, map5, map6,map7,map8,map9,map10, map11,map12, 
                           map13, map14, map15, nrow=3, ncol=5)
```

# Create neighbors 
Creating queen contiguity neighbors among each city (spatial islands)
```{r neighbors}

#create neighbors GA
gnb <- poly2nb(full_data_georgia)
g_listw <- nb2listw(gnb, style = 'W')

#create neighbors atlanta
anb <- poly2nb(full_data_atlanta)
a_listw <- nb2listw(anb, style = 'W')

#create neighbors macon
mnb <- poly2nb(full_data_macon)
m_listw <- nb2listw(mnb, style = 'W')

#create neighbors columbus
cnb <- poly2nb(full_data_columbus)
c_listw <- nb2listw(cnb, style = 'W')

#create neighbors augusta
aanb <- poly2nb(full_data_augusta)
aa_listw <- nb2listw(aanb, style = 'W')

#create neighbors columbus
snb <- poly2nb(full_data_savannah)
s_listw <- nb2listw(snb, style = 'W')

```

# Aspatial regression: Georgia
Determining whether there is evidence of *any* clustering as justification for subsequent spatial analyses.
```{r morans I GA}
#specify dependent variables for lm 
outcomes<- c("RESP", "DSLPM", "CANCER")

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_georgia)
assign(model,m)
}

#RESP
summary(model1)
#sig and positively associated 
AIC(model1)

summary(model2)
#not associated 

summary(model3)
#sig and positively associated 

#RESP
lm.morantest(model1, listw = g_listw, zero.policy = T)
#morans: 0.933920024     , p: <2.2 e-16

#DSLPM
lm.morantest(model2, listw = g_listw, zero.policy = T)
#morans:  0.897474442     , p: <2.2 e-16

#CANCER
lm.morantest(model3, listw = g_listw, zero.policy = T)
#morans:  0.933282563     , p: <2.2 e-16

```
#Spatial regression: Spatial Durbin Model (SDM) among all cities 
```{r SDM GA}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_georgia, listw = g_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```


# Aspatial regression: Atlanta
```{r morans I ATL}

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_atlanta)
assign(model,m)
}

summary(model1)
#sig and positively associated 
AIC(model1)

summary(model2)
#non-sig and positively associated 

summary(model3)
#sig and positively associated 

#RESP
lm.morantest(model1, listw = a_listw, zero.policy = T)
#morans: 0.710699601, p: <2.2 e-16

#DSLPM
lm.morantest(model2, listw = a_listw, zero.policy = T)
#morans:  0.810299413, p: <2.2 e-16

#CANCER
lm.morantest(model3, listw = a_listw, zero.policy = T)
#morans:  0.684958693, p: <2.2 e-16

```

#Spatial regression: SDM in Atlanta
```{r SDM ATL}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_atlanta, listw = a_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```

# Aspatial regression: Macon
```{r morans I macon}

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_macon)
assign(model,m)
}

summary(model1)
#not associated
AIC(model1)

summary(model2)
#not associated

summary(model3)
#not associated

#RESP
lm.morantest(model1, listw = m_listw, zero.policy = T)
#morans: 0.36166334 , p: 0.006558

#DSLPM
lm.morantest(model2, listw = m_listw, zero.policy = T)
#morans:  0.41996037, p: 0.002542

#CANCER
lm.morantest(model3, listw = m_listw, zero.policy = T)
#morans:  0.33689573    , p:  0.009534

```
#Spatial regression: SDM in Macon
```{r SDM macon}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_macon, listw = m_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```

# Aspatial regression: Columbus
```{r morans I columbus}

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_columbus)
assign(model,m)
}

summary(model1)
#not associated
AIC(model1)

summary(model2)
#not associated

summary(model3)
#not associated

#RESP
lm.morantest(model1, listw = c_listw, zero.policy = T)
#morans: 0.19419672 , p: 0.01714

#DSLPM
lm.morantest(model2, listw = c_listw, zero.policy = T)
#morans:  0.50627630, p: 1.037e-05

#CANCER
lm.morantest(model3, listw = c_listw, zero.policy = T)
#morans:  0.17682483     , p:  0.02288

```

#Spatial regression: SDM in Columbus
```{r SDM columbus}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_columbus, listw = c_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```
# Aspatial regression: Augusta
```{r morans I augusta}

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_augusta)
assign(model,m)
}

summary(model1)
#not associated
AIC(model1)

summary(model2)
#not associated

summary(model3)
#sig and positively associated

#RESP
lm.morantest(model1, listw = aa_listw, zero.policy = T)
#morans: 0.26835958  , p: 0.003272

#DSLPM
lm.morantest(model2, listw = aa_listw, zero.policy = T)
#morans:  0.36258233, p: 0.0003841

#CANCER
lm.morantest(model3, listw = aa_listw, zero.policy = T)
#morans:  0.26876071     , p:  0.003245

```

#Spatial regression: SDM in Augusta
```{r SDM augusta}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_augusta, listw = aa_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```

# Aspatial regression: Savannah
```{r morans I savannah}

#for loop: 
for(i in 1:length(outcomes)){
model <- paste("model",i, sep="")
m <- lm(as.formula(paste(outcomes[i],"~ HRS10")), data=full_data_savannah)
assign(model,m)
}

summary(model1)
#not associated
AIC(model1)

summary(model2)
#not associated

summary(model3)
#sig and positively associated

#RESP
lm.morantest(model1, listw = s_listw, zero.policy = T)
#morans: 0.55149243  , p: 9.7e-07

#DSLPM
lm.morantest(model2, listw = s_listw, zero.policy = T)
#morans:  0.73040748, p: 4.246e-10

#CANCER
lm.morantest(model3, listw = s_listw, zero.policy = T)
#morans:  0.42231295     , p:  8.289e-05

```

#Spatial regression: SDM in Savannah
```{r SDM savannah}

#for loop: 
for(i in 1:length(outcomes)){
SDMmodel <- paste("SDMmodel",i, sep="")
m <- lagsarlm(as.formula(paste(outcomes[i],"~ HRS10")), 
              data=full_data_savannah, listw = s_listw, Durbin=T)
assign(SDMmodel,m)
}

#RESP
summary(SDMmodel1)

#DSLPM
summary(SDMmodel2)

#CANCER
summary(SDMmodel3)

```


